---
all_doc:
  pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
output:
  html_vignette:
    toc: yes
    toc_depth: 2
  html_document:
    theme: flatly
    highlight: tango
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
  pdf_document:
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
vignette: |
  %\VignetteIndexEntry{ hierarchical storage }
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
title: hierarchical storage
---

```{r setup, include = FALSE}
  library(magrittr)
  library(rlang)
  library(metayer)
```

```{r, include = FALSE}
  Sys.setenv(R_CONFIG_ACTIVE = "jupyter")

  # knitr options [ref](https://yihui.org/knitr/options/) documentation; 
  knitr::opts_chunk$set(
    metayer_hook = TRUE,
    comment = "",
    fig.path = "assets/knitr/",
    
    # c.f., _pkgdown \ templates \ includes
    class.output = c("sourceCode", "rOutput")
  )

  knitr::opts_knit$set(
    root.dir = here::here("vignettes")
  )

  knitr::knit_hooks$set(
    metayer_hook = knitr_metayer_hook
  )
```

## globally available storage

```{r}
# this will create nested environments within in the global environment
rm.all()
local({
  order_id = "D1234"
  order <- storage_env("metayer", "menu", "dinner", order_id)
  order$chicken = TRUE

  meals <- storage_env("metayer", "menu")
  
  # assert that things are organized as we expect
  stopifnot(
    identical(order, meals$dinner[["D1234"]])
  )
  
  # return the order, an environment, cast as a list
  order %>%
    as.list()
})
```

`storage_env` has a side effect, and this is by design. It creates a
`.storage` container in the global environment. Subsequent calls to
`storage_env` expect to have access to this hierarchical storage
container.

```{r}
# examine objects in the global enviroment
ls(all.names = TRUE)
```

```{r}
# access the order directly
.storage$metayer$menu$dinner$D1234 %>%
  as.list()
```

```{r}
# access the order through the api
storage_env("metayer", "menu", "dinner")$D1234 %>%
  as.list()
```

## locally scoped storage

It's worth noting that the storage doesn't have to live in the global
namespace.

```{r}
# Clean up the global environment, then create a 'lunch' container in the local environment.
rm.all()
local({
  order <- storage_env("metayer", "menu", "lunch", "L0222", .store = "lunch", .envir = current_env())
  order$burger = TRUE  

  # display a list of defined objects
  ls(all.names = TRUE) %>%
    capture.output() %>%
    cat()

  order %>% 
    as.list()
})
```

Here, `storage_env` is used in a way where the global environment is not
polluted.

```{r}
ls(all.names = TRUE) %>%
  is_empty()
```
