---
all_doc:
  pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
output:
  html_vignette:
    toc: yes
    toc_depth: 2
  html_document:
    theme: flatly
    highlight: tango
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
  pdf_document:
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
vignette: |
  %\VignetteIndexEntry{ logger integration }
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
title: logger integration
---

```{r setup, include = FALSE}
  library(magrittr)
  library(rlang)
  library(metayer)
```

```{r, include = FALSE}
  Sys.setenv(R_CONFIG_ACTIVE = "jupyter")

  # knitr options [ref](https://yihui.org/knitr/options/) documentation; 
  knitr::opts_chunk$set(
    metayer_hook = TRUE,
    comment = "",
    fig.path = "assets/knitr/",
    
    # c.f., _pkgdown \ templates \ includes
    class.output = c("sourceCode", "rOutput")
  )

  knitr::opts_knit$set(
    root.dir = here::here("vignettes")
  )

  knitr::knit_hooks$set(
    metayer_hook = knitr_metayer_hook
  )
```

## logging

As code shifts from prototype to production, logging becomes important.
Given the ubiquity of the
[cli](https://cli.r-lib.org/reference/index.html) and
[logger](https://daroczig.github.io/logger/) packages, it would be
helpful to have the option of routing cli output directly into the logs.
More generally, any code that generates conditions--messages, warnings,
and errors--should be easily loggable as well. To address this concern,
metayer provides a `with_logger` function and uses it to wrap many of
the cli functions.

```{r}
## tl;dr

cli_alert("use cli methods to route output into the logs") %>%
  suppressMessages()

tryCatch(
  cli_abort("abort: an error"),
  error = function(m) {
    log_error("error: handled")
    invisible(NULL)
  }
)

msg <- with_logger({
  message("a logged message")
}) %>%
  catch_cnd()
```

## with_logger

`with_logger` provides explicit logging. Internally, this uses
`withCallingHandlers`, so conditions are processed with a log entry
before being propagated to the next handler. This results in the
visibility of both the log entry as well as the message.

```{r}
# explicity logging
with_logger({
  message("a logged message")
})
```

## cli wrappers

metayer uses some metadata magic to wrap cli functions. This imbues them
with logging functionality.

metayer also utilizes the `cli.default_handler` option provided by the
cli package. This upgrades glue semantics to generate visible NULL
values and is useful when debugging.

We consider the available variants of `cli_alert_warning`, starting with
cli's default behavior.

```{r}
# default behavior of cli_alert_warning
withr::with_options(
  list(cli.default_handler = NULL), 
  cli::cli_alert_warning("a message without a visible null: '{NULL}'")          
)
```

## cli.default_handler - visible nulls

When using the default metayer handler--`cli_nullity_handler`--we see
that NULL values are handled differently. We can also customize this
behavior.

```{r}
# visible null, default
cli::cli_alert_warning("a message with a default null: '{NULL}'") 

# visible null, customized
withr::with_options(
  list(mty.cli_null = "NULL"),
  cli::cli_alert_warning("a message with a custom null: '{NULL}'") 
)
```

## explicitly logged cli functions

If we were willing to, we could apply `wrap_logger` to every cli call.

```{r}
# logging
with_logger(
  cli::cli_alert_warning("a logged message from 'cli::cli_alert_warning'")
)
```

It's odd to see a function named `cli_alert_warning` producing a log
with an *INFO* threshold. This is easily amended. To ensure that the
logger level is aligned with the naming implications of the cli
function, use the `level` parameter. Now `cli_alert_warning` is
producing a *WARN* entry in the logs.

```{r}
# logging with log level
with_logger(
  cli::cli_alert_warning("a logged warning with 'cli::cli_alert_warning'"),
  level = logger::WARN
)
```

Finally, it's worth mentioning the double output above. This is because
`cli::cli_alert_warning` bubbles up a message condition, and
`with_logger` processes it along the way. The second output above is
that message, delivered to the user. We can deal with this using an even
more verbose expression. Now we'll only see what's propagated to the
logs.

```{r}
# fix the double messaging
tryCatch(
  {
    with_logger(
      cli::cli_alert_warning("a logged warning with 'cli::cli_alert_warning'"),
      level = logger::WARN
    )
  },
  message = function(msg) invisible(NULL)
)
```

## wrapped cli functions

metayer provides wrapped versions of cli functions. This recovers the
usual verbosity of the original cli package but without the pain
previously demonstrated.

```{r}
# implicit logging and log levels
cli_alert_warning("a logged warning with 'cli_alert_warning'")
```

## namespace context

So far, all of the logging has been to the global namespace. In an
article context, this is as it should be. However, logs originating
elsewhere should reflect the point of origin.

With `log_info` and friends, one can explicitly specify a namespace.

```{r}
# explicit custom namespace
log_warn("a warning with 'log_warn' in a custom namespace", namespace = "custom")
```

To showcase implicit namespace behavior, we'll create two functions in
the global namespace. We'll also put copies of these functions in a
namespace anchored to metayer. This is intended to mimic the scenario
where these functions would have been formally defined in the package.

```{r}
# two functions in the global namespace
foo <- function() log_info("foo called `log_info`")
bar <- function() cli_alert("bar called `cli_alert`")
```

```{r}
# global functions produce 'global' namespace
foo()
bar()
```

In the previous output, note that `foo`--inherently, `log_info`--doesn't
generate a condition; it shows up only once.

`bar`, on the other hand, calls `cli_alert` which generates a condition.
Hence that output shows up twice: once as a log entry, processed as it
happens, and a second time, now as an unhandled condition that has
bubbled up to the top level. This is the same behavior we observed
previously.

Next, we'll put `foo` and `bar` into the metayer namespace and observe
that they now display a "metayer" namespace.

```{r}
# two functions "above" the metayer namespace
ns <- new_environment(parent = getNamespace("metayer"))
ns[["foo"]] <- set_env(foo, ns)
ns[["bar"]] <- set_env(bar, ns)
```

```{r}
# packaged functions produce 'metayer' namespace
ns$foo()
ns$bar()
```

## glue semantics

The wrapped functions are designed to replicate the usual glue semantics
found in the cli and logger packages.

```{r}
# glue semantics for `cli_alert_info` (with visible NULLs)
v1 = "foo"
v2 = NULL
v3 = "bar"

cli_alert_info("{v1} {v2} {v3}")
```

```{r}
# glue semantics for log_info
log_info("a log: {v1} {v3}")
```

However, just like their analogues in the logger package, `log_info` and
friends won't handle NULL values.

```{r}
# null values were never allowed here!
logger::log_info("a log: {v1} {v2} {v3}")
```

```{r}
# and so we don't expect them to be available here, either.
log_info("a log: {v1} {v2} {v3}")
```
