---
all_doc:
  pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
output:
  html_vignette:
    toc: yes
    toc_depth: 2
  html_document:
    theme: flatly
    highlight: tango
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
  pdf_document:
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
vignette: |
  %\VignetteIndexEntry{ getting started }
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
title: getting started
---

```{r setup, include = FALSE}
  library(magrittr)
  library(rlang)
  library(metayer)
```

```{r, include = FALSE}
  Sys.setenv(R_CONFIG_ACTIVE = "jupyter")

  # knitr options [ref](https://yihui.org/knitr/options/) documentation; 
  knitr::opts_chunk$set(
    metayer_hook = TRUE,
    comment = "",
    fig.path = "assets/knitr/",
    
    # c.f., _pkgdown \ templates \ includes
    class.output = c("sourceCode", "rOutput")
  )

  knitr::opts_knit$set(
    root.dir = here::here("vignettes")
  )

  knitr::knit_hooks$set(
    metayer_hook = knitr_metayer_hook
  )
```

## publishing context

The context in which a figure is created often impacts its aesthetics,
and each publishing context may require separate options or
configuration settings to obtain a uniform output. `metayer::sure_plot`
provides a DRY solution.

The [authoring vignette](./authoring.html) describes patterns to
uniformize output at each stage of the publishing pipeline.

```{r}
sure_plot({
    plot(pcd$x, pcd$y)
  },
  height = 5,
  width = 5,
  par_opts = list(bg = "#f1d08f")
)
```

## cli / logger integration

As code shifts from prototype to production, logging becomes important.
Given the ubiquity of the [cli](https://cli.r-lib.org/index.html) and
[logger](https://daroczig.github.io/logger/index.html) packages, it
would be nice if messages that were originally headed for an interactive
terminal could be seamlessly integrated into the logs.

The [logger integration vignette](./logger_integration.html) shows the
details of how metayer makes this work.

```{r}
# contextualized cli calls
cli_alert("do what you love!") %>% 
  suppressMessages()
```

## global hierarchical storage

metayer implements a global hierarchical storage container which is
simple to use. The [hierarchical storage
vignette](./hierarchical_storage.html) does a deeper dive.

```{r}
# pluck-like: create a persistent, hierarchically indexed environment
rm.all()
order_env <- local({
  order_id <- "D1234"
  storage_env("metayer", "menu", "dinner", order_id)
})
identical(.storage$metayer$menu$dinner[["D1234"]], order_env)
```

## metaprogramming helpers

metayer provides two methods for easing metaprogramming tasks,
`wrapped_factory` and `with_monkey_patch`. Specifics are available in
the [metaprogramming vignette](./metaprogramming.html).

## configurability

metayer is designed to be incorporated into future projects. As such, it
aims to be configurable. Details are in the [configuration
vignette](./metayer-config.html).

## miscellanea

Some other useful metayer utilities can be found in the [miscellanea
vignette](miscellanea.html).
