---
all_doc:
  pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
output:
  html_vignette:
    toc: yes
    toc_depth: 2
  html_document:
    theme: flatly
    highlight: tango
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
  pdf_document:
    pandoc_args: --resource-path=/home/dnlennon/Workspace/repos/metayer/vignettes
vignette: |
  %\VignetteIndexEntry{ metaprogramming }
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
title: metaprogramming
---

```{r setup, include = FALSE}
  library(magrittr)
  library(rlang)
  library(metayer)
```

```{r, include = FALSE}
  Sys.setenv(R_CONFIG_ACTIVE = "jupyter")

  # knitr options [ref](https://yihui.org/knitr/options/) documentation; 
  knitr::opts_chunk$set(
    metayer_hook = TRUE,
    comment = "",
    fig.path = "assets/knitr/",
    
    # c.f., _pkgdown \ templates \ includes
    class.output = c("sourceCode", "rOutput")
  )

  knitr::opts_knit$set(
    root.dir = here::here("vignettes")
  )

  knitr::knit_hooks$set(
    metayer_hook = knitr_metayer_hook
  )
```

metayer provides two methods for easing metaprogramming tasks,
`wrapped_factory` and `with_monkey_patch`.

## wrapped_factory

It's easier to start with `wrapped_factory` which will allow us to wrap
existing functions. One feature of `wrapped_factory` is that it
replicates the wrapped function's signature, including default values
and parameters that would otherwise be lazily evaluated.

We note that wrappers should adopt the form `function(cmd, args, ...)`.
`wrapped_factory` will substitute `cmd` and `args` in the wrapped
function. `...` may be replaced with key-value pairs which will be
available in the wrapper scope.

An example should be illustrative. Let's define a wrapper that adds
printf debugging to an existing function.

```{r}
# a simple "debug" wrapper
debug_wrapper <- function(cmd, args, label = NULL) { 
  
  # emit debugging information
  sprintf(">>> called '%s'\n", label) %>%
    cat(file = stdout())

  # call the original function
  do.call(cmd, args)
}
```

We'll apply this wrapper to a simple function:

```{r}
# a very simple function
sum <- function(x, y) x + y
```

Now, invoke the `wrapped_factory` machinery to produce a wrapped
function. This will print the debugging information on stdout and the
return the result.

```{r}
dbg_sum <- wrapped_factory("sum", debug_wrapper, label = "..sum..")
dbg_sum(1, 2)
```

It's instructive to inspect the structure of `dbg_sum`. It has the same
function signature as the original function; the body of the function is
`debug_wrapper` with adapted `cmd` and `args`. Specifically, `cmd` has
been replaced with `sum`, and `args` has been replaced with a
symbol-mapped list. `wrapped_factory` makes these changes with the
`substitute` function, and these replacements will happen wherever these
symbols are found in the wrapper code.

```{r}
dbg_sum
```

## with_monkey_patch

`with_monkey_patch` does roughly the same thing as `wrapped_factory`,
but it is applied to functions defined in a namespace and only
temporarily. Changes will be restored on exit.

The following is a simple example. Consider the behavior of
"base::Sys.time": it captures the local time zone in its output.

```{r}
# e.g., "PDT" in Seattle.
Sys.time()
```

That means that a function, like the one below, is inherently broken.
It's an inflexible implementation, only meaningful in a single, fixed
timezone determined apriori.

```{r}
# hmm...
get_time_string <- function() {
  Sys.time() %>% as.character()
}
get_time_string()
```

`with_monkey_patch` allows us to to fix `get_time_string` indirectly. It
does this by temporarily modifying the behavior of `base::Sys.time` in a
scoped block of client code. Here, calls to `base::Sys.time` will return
UTC.

```{r}
# changes the behavior of base::Sys.time so that it returns UTC
with_monkey_patch(
  "base::Sys.time",
  # 
  wrapper = function(cmd, args, func) {    
    t <- do.call(func, args)
    .POSIXct(t, "UTC")
  },
  {
    # scoped block: (nested) calls to Sys.time only take effect here
    get_time_string()
  }
)
```

```{r}
# check that the original behavior is restored
get_time_string()
```

### a real example

While the above would hopefully never happen in production code, there
are examples where the monkey patch machinery is useful for modifying an
object created deep in a call stack.

metayer initially used this monkey patching pattern to affix knitr hooks
to rmarkdown documents. Specifically, `pkgdown::build_article` invokes
`rmarkdown::html_document`, but does so deep in the call stack and with
limited configurability. Our scenario requires a knitr hook to be
installed, and while it is trivial to modify an html_document object for
this purpose, access to this object wasn't exposed through the pkgdown
API. One proposal was to monkey patch the behavior of
`rmarkdown::html_document` so that it included the post-processing steps
that were required.

Ultimately, the monkey patching approach for `rmarkdown::html_document`
was refactored away, but, for a time, it provided a viable solution to
the original problem.
