---
title: "configuration"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: sentence
vignette: >
  %\VignetteIndexEntry{configuration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
  library(metayer)
  library(magrittr)
```

```{r, include = FALSE}
  knitr::opts_chunk$set(
    chunk.comment = "",
    fig.path = "assets/knitr/",
    class.output = c("sourceCode", "rOutput")
  )

  knitr::opts_knit$set(
    root.dir = here::here("vignettes")
  )
```

When metayer is loaded, it populates several options and sets up logging.  Much of this can be configured by making changes to config.yml.

## configurability

Package configuration is specified in config.yml, which is designed to mimic the behavior of the [config](https://rstudio.github.io/config/reference/index.html) package.  As such, the top-level sections--default, testing, and jupyter--are intended to correspond to an execution / operational mode.  Our extension, `config_get`, enables YAML references and provides a few default handlers for custom YAML tags.

```{r results='asis', echo=FALSE}
yaml_content <- get_raw_yaml("default")
sprintf("```yaml\n%s\n```", yaml_content) %>% cat()
```

## custom YAML tags

Note that there are custom YAML tags available, `!optenv` and `!with_env`.

### !optenv

optenv is used to evaluate simple R expressions, like `logger::INFO`; the idea is to simplify the specification of objects or functions.

### !with_env

with_env allows strings to contain environment variables.

## config.yml

### logger

#### format

A string specification for the logging format passed to logger::log_format

#### logfile

The location of a log file.  This will become the log_appender at namespace global and index one.

#### appenders

A list of appenders that will populate subsequent indexes of the global namespace.

#### threshold

The log threshold, e.g., logger::INFO

### options

#### cli.&ast;

[cli package options](https://cli.r-lib.org/reference/cli-config.html)

#### cli.default_handler

We set this to `cli_nullity_handler` in order to enjoy null visibility in cli output.

#### cli.ansi

Useful in testing to ensure that output doesn't have escape codes.

#### cli.unicode

Useful in testing to ensure that output doesn't have unusual unicode characters.

#### cli.num_colors

Useful in testing to ensure that output doesn't have escape codes.

#### cli.condition_unicode_bullets

Useful in testing to ensure that output doesn't have unusual unicode characters.


#### knitr.chunk.&ast;

[knitr chunk options](https://yihui.org/knitr/options/#chunk-options)

#### knitr.chunk.dev

  The default device used when knitting.

#### knitr.chunk.comment

A cosmetic change for  R output.

#### knitr.chunk.class.output

&nbsp;

#### knitr.chunk.error

An aesthetic choice for R output.

A cosmetic change for imbuing R output with class tags suitable for CSS.

#### knitr.chunk.metayer_hook

A custom knitr options.  Setting this enables [sure_plot](./reference/sure_plot.html) which requires a custom knitr hook (specified in .onLoad).

#### knitr.package.&ast;

[knitr package options](https://yihui.org/knitr/options/#package-options)

#### knitr.package.root.dir

The directory in which knitr will expect to find references.  pkgdown and rmarkdown behavior can be quite sensitive to this.

#### mty.cli_null

The substitution made for NULL values in cli output.

#### mty.fig.&ast;

`with_grdev` default parameters.  See the function [reference](reference/with_grdev.html).

#### uuid.generator

When testing, set this to test_mty_uuid for reproducible uuid-like output.

#### uuid.generator.seed

A seed value for the uuid.generator.

### display

#### ippx, ippy

Inches per pixel.  Not currently used.  This would be helpful for creating interactive devices--e.g. X11--with dimensions specified in pixels rather than inches.

### rdoc_config

This section serves as the default YAML header for new jupyter notebooks.  With this, one really only needs to specify a title when creating new documents.

#### output

The usual jupyter YAML content.

```{r results='asis', echo=FALSE}
yaml_content <- get_raw_yaml("default", "rdoc_config", "output")
sprintf("```yaml\n%s\n```", yaml_content) %>% cat()
```

### rmall_exclusions

A list of functions to blacklist when running `rm.all`.
